---
title: "STA410 A1"
output: pdf_document
---

Name: Harold Hyun Woo Lee
\newline

R coding part of the assignment \newline
##Question1)
```{r}
#fwht2d
fwht2d <- function(x) {
             h <- 1
             len <- ncol(x)
             while (h < len) {
                for (i in seq(1,len,by=h*2)) {
                   for (j in seq(i,i+h-1)) {
                     a <- x[,j]
                     b <- x[,j+h]
                     x[,j] <- a + b
                     x[,j+h] <- a - b
                     }
                }
             h <- 2*h
             }
             h <- 1
             len <- nrow(x)
             while (h < len) {
                for (i in seq(1,len,by=h*2)) {
                   for (j in seq(i,i+h-1)) {
                     a <- x[j,]
                     b <- x[j+h,]
                     x[j,] <- a + b
                     x[j+h,] <- a - b
                     }
                }
             h <- 2*h
             }
             x
             }

```
#1b&c)
```{r}
#1b Hard and soft thresholding)
hard <- function(x, lambda){
  #hard is a function that takes in a single element and a lambda
  if (abs(x) >= lambda){
    
    return(x)
  }
  else{
    return(0)
  }
}

soft <- function(x, lambda){
  #soft is a function that takes in a single element and a lambda
  if ((abs(x) - lambda)>=0){
    return(sign(x)*(abs(x) - lambda))
  }
  else{
    return(0)
  }
}

#Image is the original image.
design <- matrix(scan("design.txt"),ncol=256,byrow=T)
colours <- grey(seq(0,1,length=256))
image(design, axes=F, col=colours)

par(mfrow=c(2,2))

#I will make 4 denoised images using hard threshold
xhat <- fwht2d(design) # x is a 2^k x 2^k matrix
xhat_2 <- fwht2d(design)
xhat_3 <- fwht2d(design)
xhat_4 <- fwht2d(design)

#First denoised image using Hard
#Every element in the first column will be thresholded against lambda of 0
for (x in (1:256)){
  xhat[x] <- hard(xhat[x], 0)
}
#Then elements in other columns will be thresholded against lambda of 150
for (x in (257:length(design))){
  xhat[x] <- hard(xhat[x], 150)
}


#Second denoised image using Hard
#Every element in the first two column will be thresholded against lambda of 0
for (x in (1:512)){
  xhat_2[x] <- hard(xhat_2[x], 0)
}
#Then elements in other columns will be thresholded against lambda of 150
for (x in (573:length(design))){
  xhat_2[x] <- hard(xhat_2[x], 150)
}


#Third denoised image using Hard
#Every element in the first column will be thresholded against lambda of 0
for (x in (1:256)){
  xhat_3[x] <- hard(xhat_3[x], 0)
}
#Then elements in other columns will be thresholded against lambda of 300
for (x in (257:length(design))){
  xhat_3[x] <- hard(xhat_3[x], 300)
}


#Fourth denoised image using Hard
#Every element in the first column will be thresholded against lambda of 0
for (x in (1:256)){
  xhat_4[x] <- hard(xhat_4[x], 0)
}
#Then elements in other columns will be thresholded against lambda of 75
for (x in (257:length(design))){
  xhat_4[x] <- hard(xhat_4[x], 75)
}


xhat_inverse <- fwht2d(xhat)/ncol(xhat)^2 # inverse transform
xhat_2_inverse <- fwht2d(xhat_2)/ncol(xhat_2)^2
xhat_3_inverse <- fwht2d(xhat_3)/ncol(xhat_3)^2
xhat_4_inverse <- fwht2d(xhat_4)/ncol(xhat_4)^2

#These are the images when Hard threshold is used.
image(xhat_inverse, axes=F, col=colours)
image(xhat_2_inverse, axes=F, col=colours)
image(xhat_3_inverse, axes=F, col=colours)
image(xhat_4_inverse, axes=F, col=colours)
```
Note that the above 4 graphs used hard thresholding method. 
\newline
Top left: first column lambda is 0, other columns lambda is 150 \newline
Top right: first 2 column lambda is 0, other columns lambda is 150 \newline
Bottom left: first column lambda is 0, other columns lambda is 300 \newline
Bottom right: first column lambda is 0, other columns lambda is 75 \newline
\newline
We can see that using hard threshold with lambda of around 150 is enough to denoise the image just right.
Anything more or less than lambda value of 150, we get to see more blurriness in the denoised image.

```{r}
par(mfrow=c(2,2))

#I will make 4 denoised images using soft threshold now
soft_1 <- fwht2d(design) # x is a 2^k x 2^k matrix
soft_2 <- fwht2d(design)
soft_3 <- fwht2d(design)
soft_4 <- fwht2d(design)

#First denoised image using soft
#Every element in the first column will be thresholded against lambda of 0
for (x in (1:256)){
  soft_1[x] <- soft(soft_1[x], 0)
}
#Then elements in other columns will be thresholded against lambda of 150
for (x in (257:length(design))){
  soft_1[x] <- soft(soft_1[x], 150)
}


#Second denoised image using soft
#Every element in the first two column will be thresholded against lambda of 0
for (x in (1:512)){
  soft_2[x] <- soft(soft_2[x], 0)
}
#Then elements in other columns will be thresholded against lambda of 150
for (x in (573:length(design))){
  soft_2[x] <- soft(soft_2[x], 150)
}


#Third denoised image using soft
#Every element in the first column will be thresholded against lambda of 0
for (x in (1:256)){
  soft_3[x] <- soft(soft_3[x], 0)
}
#Then elements in other columns will be thresholded against lambda of 300
for (x in (257:length(design))){
  soft_3[x] <- soft(soft_3[x], 300)
}


#Fourth denoised image using soft
#Every element in the first column will be thresholded against lambda of 0
for (x in (1:256)){
  soft_4[x] <- soft(soft_4[x], 0)
}
#Then elements in other columns will be thresholded against lambda of 75
for (x in (257:length(design))){
  soft_4[x] <- soft(soft_4[x], 75)
}


soft_1_inverse <- fwht2d(soft_1)/ncol(soft_1)^2 # inverse transform
soft_2_inverse <- fwht2d(soft_2)/ncol(soft_2)^2
soft_3_inverse <- fwht2d(soft_3)/ncol(soft_3)^2
soft_4_inverse <- fwht2d(soft_4)/ncol(soft_4)^2

#These are the images when Hard threshold is used.
image(soft_1_inverse, axes=F, col=colours)
image(soft_2_inverse, axes=F, col=colours)
image(soft_3_inverse, axes=F, col=colours)
image(soft_4_inverse, axes=F, col=colours)
```
Note that the above 4 graphs used soft thresholding method. 
\newline
Top left: first column lambda is 0, other columns lambda is 150 \newline
Top right: first 2 column lambda is 0, other columns lambda is 150 \newline
Bottom left: first column lambda is 0, other columns lambda is 300 \newline
Bottom right: first column lambda is 0, other columns lambda is 75 \newline
\newline
We can see that using soft threshold with lambda of around 75 is enough to denoise the image just right.
Anything more or less than lambda value of 75, we get to see more blurriness in the denoised image.\newline

I will now try the same thing but instead of hard threshold, using soft threshold. \newline

##Question 2d)
```{r}
#Define our variables
theta = 0.9
e = 10^(-5)

#Let us first find the M value
#compute the value of t such that Phi(t) = 1/theta
upper_t = ((1/theta)*(2^10))^(1/10) - 1
lower_t = 1
#Hence (lower_t, upper_t) is the interval for Phi(t) such that 1<Phi(t)<1/theta
#So let's take a discrete set of points T such that T = {1< t_1, t_2, - - - , t_k}
#Let t_k = 1.02
t<- seq(1, 1.02, len=1000)

#Define phi function
Phi <- function(t){
  ((1+t)^10)/(2^10)
}

#Let's compute for M
find_M <- function(t){
  M <- min((log(1 - theta) - log(1-theta*Phi(t)) - log(e))/log(t))
}
#I'll use floor function here to keep M as an integer
M <- floor(find_M(t))
print(M)

x_value <- seq(1, 1.02, len=1000)
y_value <- (log(1 - theta) - log(1-theta*Phi(x_value)) - log(e))/log(x_value)
x_new <- x_value[950:1000]
y_new <- y_value[950:1000]
plot(x_new,y_new, xlab = "Theta(t)", ylab = "M", main = "Plot of M value for a given Theta(t)")

#using the plot we can tell that M=724 is the minimum value.
```

```{r}
#2d
#Define distribution of{X_i}
p <- function(x){
  choose(10, x)*(1/2)^10
}
#Define DFT of {p(x) :x=0, ... , M-1}
phat <- function(j){
  x = seq(0, (M-1), 1)
  i = complex(real=0, imaginary=1)
  sum(exp(-2*pi*i*(j/M)*x)*p(x))
}

#Define g function
g <- function(x){
  (1-theta)/(1-theta*x)
}
#Define inverse FFT
PS <- function(s){
  result = 0
  i = complex(real=0, imaginary=1)
  for (j in (0:723)){
    var = exp(2*pi*i*(s/M)*j)*g(phat(j))
    result = result+var
  }
  return((1/M)*result)
}
```
#2d part 1)
```{r}
# Evaluate the DFT
x = c(seq(0, M-1, 1))
DFT = c()
for (var in x){
  DFT_var = phat(var)
  DFT = c(DFT, DFT_var)
}

#DFT now stores phat(j) values where j = 0, --- , M-1

#DFT
```
#2d Part 2
```{r}
# Evaluate g(phat(j))
#we already computed and stored phat(j) values in variable called DFT.
#Loop through this and calculate g(phat(j)) for j = 0, --- , M-1

gphat <- c()
for (var in DFT){
  gphat = c(gphat, g(var))
}
#gphat now stores all g(phat(j)) for j = 0, --- , M-1
#gphat
```

#2d Part 3
```{r}
S = seq(0, 723, 1)
ps <- c()
for (var in S){
  ps = c(ps, PS(var))
}
#ps
#ps is the P(S=s) for s = {0,...,M-1}
#Here using sum(ps) we can see that it totals up to 1, which is what we want.
sum(ps)

plot(c(0:(M-1)),ps,type="h",lwd=3, xlim=c(0,200), xlab = "S=s", ylab = "P(S=s)", 
     main="Probability graph for P(S=s)")
```
For the sake of graph, I only graphed up to s value of 200 as we can tell that P(S=s) approach 0. 
Nonetheless, sum of P(S=s) is 1 and this is exactly what we want.

Written part of the assignment is below.
