---
title: "STA355 A2"
author: "Harold Hyun Woo Lee"
date: "2/21/2021"
output: pdf_document
---
__Q1a)__
```{r}
dat <- scan("buffsnow.txt")

plot(density(dat, bw=5), main="Density with bandwidth 5 and 2 modes", xlab="Snowfall")
plot(density(dat, bw=4), main="Density with bandwidth 4 and 3 modes", xlab="Snowfall")

```
\newpage
 __Q1c)__ 
```{r}
kde.cv <- function(x,h) {
             n <- length(x)
             if (missing(h)) {
                r <- density(x)
                h <- r$bw/4 + 3.75*c(0:100)*r$bw/100
                }
             cv <- NULL
             for (j in h) {
                cvj <- 0
                for (i in 1:n) {
                   z <- dnorm(x[i]-x,0,sd=j)/(n-1)
                   cvj <- cvj + log(sum(z[-i]))
                   }
                cv <- c(cv,cvj/n)
                }
              r <- list(bw=h,cv=cv)
              r
              }

# Split the data into two 
dat1 <- dat[1:68]
dat2 <- dat[69:136]

# Compute bandwidth for dat1
r1 <- kde.cv(dat1)
#plot(r1$bw,r1$cv) # plot of bandwidth versus CV
r1$bw[r1$cv==max(r1$cv)] # bandwidth maximizing CV

# Compute bandwidth for dat2
r2 <- kde.cv(dat2)
#plot(r2$bw,r2$cv) # plot of bandwidth versus CV
r2$bw[r2$cv==max(r2$cv)] # bandwidth maximizing CV


plot(density(dat1, bw=7.54), main="Density for first 68 seasons", xlab="Snowfall")
plot(density(dat2, bw=17.15), main="Density for last 68 seasons", xlab="Snowfall")
```
The estimated density for the first 68 seasons is approximately 0.02, and the estimated density for the last 68 seasons is approximately 0.012.
We see that the bandwidths are 7.54 and 17.15 respectively. The later half of the data is somewhat more dispersed and has a larger bandwidth value.
In addition, we see that the density for the first 68 seasons is much higher than the density for the later half. \newpage



__Q2b)__
```{r}
x1 <- seq(0.01, 1, by=0.01)
plot(x1, gamma(1/x1 + 1)^x1, main="Limit of MPS(F) as alpha approaches 0", 
     xlab = "alpha", ylab="MPS(F)")

x2 <- seq(5000, 70000, by = 1000)
plot(x2, gamma(1/x2 + 1)^x2, main="Limit of MPS(F) as alpha approaches inf", 
     xlab = "alpha", ylab="MPS(F)")
abline(h = 0.5614661, col="red")
tail(gamma(1/x2 + 1)^x2, 1)
```
The 2 graphs above shows that as $\alpha \rightarrow 0$, $MPS(F) \rightarrow \inf$. \
In addition, as $\alpha \rightarrow \inf$, $MPS(F) \rightarrow 0.5614$. \newpage

__Q2c)__
```{r}
income <- scan("incomes.txt")


x_bar <- mean(income)
count = c()
for (val in income){
  if (val < x_bar){
    count = c(count, 1)
  }
  if (val > x_bar){
    count = c(count, 0)
  }
}
MPS_hat = sum(count)/200


loo <- NULL
for (i in 1:200){
  loo <- c(loo, mean(count[-i]))
}
sehat <- sqrt(199*sum((loo-mean(loo))^2)/200)
sehat



MPS_hat
```
Here we see that $\hat{MPS(F)} = 0.835$ and the estimate of the standard error using jackknife is 0.026 \newpage

___2d)___
```{r}

sigma_2 = var(log(income))

var_MPS = sigma_2/(16*pi) * exp(-sigma_2/4)

SE = sqrt(var_MPS)/sqrt(200)
SE
```
The estimate of the standard error of $\psi(\hat{\sigma})$ for the income data is about 0.0120. This is smaller than the estimate found in part C.
