---
title: "STA355 A1"
author: "Harold Hyun Woo Lee"
date: "1/27/2021"
output: pdf_document
---
___Question 1b)___
```{r}
library(fdrtool)
tau <- seq(from=0, to=1, by=0.01)
tau <- tau[2:100]
r <- (tau*(1-tau))/(((qhalfnorm(tau))^2)*((dhalfnorm(qhalfnorm(tau)))^2))

plot(tau, r, xlab="Tau", ylab="Gamma Squared",pch=ifelse(tau==0.86, 19, 1),  
     col=ifelse(tau==0.86, "red", "black"), main="Tau against gamma squared")


plot(tau[77:92], r[77:92], pch=ifelse(tau[77:92]==0.86, 19, 1), 
     col=ifelse(tau[77:92]==0.86, "red", "black"), xlab="Tau", ylab="Gamma Squared", 
     main="Tau against Gamma Squared zoomed in")
# Find the minimum value of r squared
min(r)

```
Here we can see that the $\tau=0.86$ minimizes the gamma squared function at 0.77

___Question 1d)___

```{r}

dat <- scan("prob1-data.txt")


halfnormal <- function(x,tau=0.5,ylim) {
       sigma <- quantile(abs(x),probs=tau)/sqrt(qchisq(tau,1))
       n <- length(x)
       pp <- ppoints(n)
       qq <- sqrt(qchisq(pp,df=1))
# upper envelope
       upper <- sigma*(qq + 3*sqrt(pp*(1-pp))/(2*sqrt(n)*dnorm(qq)))
# lower envelope 
       lower <- sigma*(qq - 3*sqrt(pp*(1-pp))/(2*sqrt(n)*dnorm(qq)))
# add upper and lower envelopes to plot
       if (missing(ylim)) ylim <- c(0,max(c(upper,abs(x))))
       plot(qq,sort(abs(x)),
          xlab="Half Normal quantiles",ylab="ordered data",pch=20,
          ylim=ylim, col=ifelse(sort(abs(x)) > 1.2*sigma*qq, "blue", "chartreuse"))
       lines(qq,lower,lty=3,lwd=3,col="red")
       lines(qq,upper,lty=3,lwd=3,col="red")
       abline(a=0,b=sigma,lwd=3)
       }




halfnormal(dat, tau=0.5)
```
For the above graph, I edited the half normal function so that points that are above the sigma slope line by 20% are colored in blue.
The rest of the points that are more or less close to the straight line whose slope is sigma are colored in neon green.
I've given it a weight of 1.2 to color the points by blue to give a sort of leeway to those points that are close to the slope but slightly above the slope.
We can see that there are about 12 points whose means are non-zero.

___Question 2c)___
```{r}

x <- scan("kevlar.txt")
x <- sort(x)
n <- length(x)
d <- c(n:1*c(x[1], diff(x)))
plot(c(1:n)/n, cumsum(d)/sum(x), xlab="t", ylab="TTT")
abline(0,1)
```
As shown above in TTT plot for the Kevlar 373/epoxy strands data, we can see that the hazard function is roughly increasing. 
We can note that the points are mostly above the 45 degree line and roughly sketch out the concave shape from $t = 0$ to $t=0.8$. 












