---
title: "STA355 A3"
author: "Harold Hyun Woo Lee"
date: "3/9/2021"
output: pdf_document
---
### Q1c) 
```{r}
dat <- scan("bees.txt")

# Convert the data from degrees to radians
dat <- dat*pi/180
n <- length(dat)

r <- ((sum(cos(dat)))^2+(sum(sin(dat)))^2)^(1/2)
k_hat <- ((r/n)*(2-(r^2/n^2)))/(1-(r^2/n^2))

# Helper function to get valid range of K
k_range <- function(x, N, delta = c(0.05, 0.2)){
  
  # Calculate h
  h <- diff(delta)/N
  
  # Get the sequence
  thetas <- seq(from = delta[1], to = delta[2], by = h)
  
  return(list("h" = h, "theta" = thetas))
}

k_range2 <- function(x, N, del){
  
  # Calculate h
  h <- del*2/N
  
  # Get the sequence
  thetas <- seq(from = k_hat - del, to = k_hat + del, by = h)
  
  return(list("h" = h, "theta" = thetas))
}






U <- function(k, lambda, r, n){
  
  before <- (exp(-lambda*k)*besselI(r*k, 0))/(besselI(k,0))^n
  prenorm <- before/max(before)
  
  prenorm

}

denom <- function(x, lambda, k, N, h){
    n <- length(x)
    r <- ((sum(cos(x)))^2+(sum(sin(x)))^2)^(1/2)
    
    post <- U(k, lambda, r, n)
    
    mult <- c(1/2,rep(1,N-1),1/2) # multipliers for trapezoidal rule  
    norm <- sum(mult*post)/(1/h) # integral evaluated using trapezoidal rule  
    post <- post/norm # normalized posterior
    post.cdf <- cumsum(mult*post)/(1/h) # compute the posterior cdf
    lower <- max(k[post.cdf<0.025])
    upper <- min(k[post.cdf<0.975])
      
    return(list("post" = post, "post.cdf" = post.cdf, 
                "lower" = lower, "upper" = upper))
  
}


k_r <- k_range2(dat, N=1000, 0.13)$theta
h <- k_range2(dat, N=1000, 0.13)$h

test <- denom(dat, lambda = 1, k_r, N = 1000, h)

plot(k_r, test$post, main="Lambda = 1", cex=0.2)

test2 <- denom(dat, lambda = 0.1, k_r, N = 1000, h)

plot(k_r, test2$post, main="Lambda = 0.1", cex=0.2)
```
The posterior densities for two lambda values are shown above in the plots. The difference between the two lambdas is that for lambda = 1, the tail on the left side is slightly bigger than lambda = 0.1 Otherwise, they are very much similar to each other.
As for the possibility that k = 0, I believe we can't rule it out. If you look at the posterior density graph, at k = 0 we see the posterior density is almost 5. If the probability was 0 then posterior density wouldn't be so high. 



### Q2c)
```{r}
library(MASS)

set.seed(1)
var_norm <- function(n){

  # M is 500
  nrep <- 500

  # Declaring x values
  x <- c(1:n)/n

  beta <- NULL

  for (i in 1:nrep) {
    y <- rnorm(n)
    r <- lmsreg(y~x)
    beta <- c(beta,r$coef[2])
    }
  var <- var(beta)
  return(var)
}

B1_norm_var <- c(var_norm(50), var_norm(100), var_norm(500), var_norm(1000), var_norm(5000))

log_B1_norm_var <- log(B1_norm_var)

x <- c(log(50), log(100), log(500), log(1000), log(5000))

fit <- lm(log_B1_norm_var~x)
alpha = fit$coefficients[2]
-alpha
```

### Q2d)
```{r}
set.seed(1)
var_cauchy <- function(n){

  # M is 500
  nrep <- 500

  # Declaring x values
  x <- c(1:n)/n

  beta <- NULL

  for (i in 1:nrep) {
    y <- rcauchy(n)
    r <- lmsreg(y~x)
    beta <- c(beta,r$coef[2])
    }
  var <- var(beta)
  return(var)
}

B1_cau_var <- c(var_cauchy(50), var_cauchy(100), var_cauchy(500), var_cauchy(1000), var_cauchy(5000))

log_B1_cau_var <- log(B1_cau_var)

x <- c(log(50), log(100), log(500), log(1000), log(5000))

fit2 <- lm(log_B1_cau_var~x)
alpha2 = fit2$coefficients[2]
-alpha2
```

