---
title: "STA355 Midterm"
author: "Harold Hyun Woo Lee"
date: "3/3/2021"
output: pdf_document
---
___Q1)a)ii)___
```{r}
tau <- seq(from=0, to=1, by=0.001)

tau <- tau[2:1000]

r <- (tau*(1-tau))/(((qexp(tau))^2)*((dexp(qexp(tau)))^2))

plot(tau[780:810],r[780:810], main="Tau against Variance", ylab="Variance", xlab="Tau")

min(r)

tau[match(min(r), r)]
```
The value of $\tau$ that minimizes $\sigma^2(\tau)$ is $\tau = 0.797$ 

___Q1)b)___
```{r}
dat <- scan("prob1.txt")
n <- 50

Y <- c(n*dat[1])

for (i in 1:49){
  Y <- c(Y, (n-i)*(dat[i+1] - dat[i]))
}

lambda <- 50/sum(Y)
lambda


SE <- (50/(lambda^2))^(-1/2)
SE
```
Computing the MLE estimate of $\lambda$ based on the data, we get $\lambda_{MLE} = 0.405$

Computing the estimate of the standard error of the estimate in (i), we get $\hat{SE}(\hat{\lambda})$ = 0.0573

___Q2)b)___
```{r}
dat2 <- scan("prob2.txt")


theta_hat <- 1-1/(mean(dat2))*exp(mean(log(dat2)))
theta_hat


# Compute pseudo-values
pseud <- NULL

for (i in 1:200){
  xi <- dat2[-i]
  A_F <- 1-1/(mean(xi))*exp(mean(log(xi)))
  pseud <- c(pseud, 200*theta_hat - 199*A_F)
}

SE_hat <- sqrt(var(pseud)/200)
SE_hat
```


Computing the estimate of Atkinson Index based on the data, we get $\hat{A}(F) = 0.843$

Computing the estimate of the standard error of the estimate of Atkinson Index, we get $\hat{SE}(\hat{A}(F))$ = 0.0305












